---
title: "Reproducible Research: Peer Assessment 1"
author: "Saurabh"
date: "10 April 2016"
output: html_document
---

## Loading and preprocessing the data


```{r,echo=TRUE}
setwd("~/Analytics/Data Science/Coursera/5. Reproducible Research")
steps_data<-read.csv("activity.csv")
str(steps_data)
steps_data$date<-as.Date(as.character(steps_data$date))
str(steps_data)
```


## What is mean total number of steps taken per day?

```{r, echo=FALSE}
library(sqldf)
library(ggplot2)
steps_data1<-sqldf("select date, sum(steps) as num_steps from steps_data group by date")
steps_data1$num_steps<-as.numeric(steps_data1$num_steps)
g<-ggplot(steps_data1,aes(x=num_steps))
g+geom_histogram()+geom_vline(xintercept = mean(steps_data1$num_steps))
summary(steps_data1$num_steps)
```

## What is the average daily activity pattern?
Interval=835, where average number of steps is maximum

```{r,echo=TRUE}
library(sqldf)
df<-sqldf("select interval, avg(steps) as avg_steps from steps_data group by interval")
library(ggplot2)
g<-ggplot(df,aes(interval,avg_steps))
g+geom_line()
```

## Imputing missing values
Total number of missing values `sum(is.na(steps_data))`. Only steps column has missing values in the dataframe.

Adding a columns with average steps against each interval so that it can be used to replace missing values. Once added then using mutate function to modify the data.

```{r,echo=TRUE}
library(sqldf)
steps_data_2<-sqldf("select a.date, a.interval, b.avg_steps,a.steps
                       from  steps_data a left join df b on a.interval=b.interval")
library(dplyr)
steps_final<-mutate(steps_data_2, steps_full=ifelse(is.na(steps),avg_steps,steps))
steps_final<-select(steps_final,date,interval,steps_full)
```

Making a historgram and seeing the summary stats for imputed data set
```{r,echo=TRUE}
library(sqldf)
library(ggplot2)
steps_final_hist<-sqldf("select date, sum(steps_full) as Total_Steps from steps_final group by date")
g<-ggplot(steps_final_hist,aes(x=Total_Steps))
g+geom_histogram()+geom_vline(xintercept=mean(steps_final_hist$Total_Steps), color="red")
summary(steps_final_hist$Total_Steps)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}
steps_final<-mutate(steps_final, day=ifelse(weekdays(date)=="Saturday" |weekdays(date)
                                                      =="Sunday","Weekend","Weekday"))
df_final<-sqldf("select interval,day, avg(steps_full) as avg_steps from steps_final
                group by interval,day")
g<-ggplot(df_final,aes(interval,avg_steps))
g+geom_line()+facet_grid(.~day)
```
